---
title: "Exercise 2"
author: "Paddy"
date: "12 March 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Wrangling - Exercise 1: Basic Data Manipulation
In this exercise, you'll work with one of the most popular starter data sets in data science, the Titanic data set. This is a data set that records various attributes of passengers on the Titanic, including who survived and who didn’t.

## Exerise 
Using R, you’ll be handling missing values in this data set, and creating a new data set.

```{r}
library(readr)
```

### Problem 1
**0: Load the data in RStudio**
Save the data set as a CSV file called titanic_original.csv and load it in RStudio into a data frame.

```{r}
titanic3 <- read_csv("~/Springboard/Exercise problems/Exercise 2/titanic3.csv")
library(tidyr)
library(dplyr)
```


### Problem 2
**1: Port of embarkation**
The embarked column has some missing values, which are known to correspond to passengers who actually embarked at Southampton. Find the missing values and replace them with S. (Caution: Sometimes a missing value might be read into R as a blank or empty string.

```{r}
titanic <- titanic3 %>% mutate(embarked = replace(embarked, is.na(embarked), "S"))
```


### Problem 3
**2: Age **
You’ll notice that a lot of the values in the Age column are missing.
Calculate the mean of the Age column and use that value to populate the missing values


```{r}
mean_age <- mean (titanic$age, na.rm = TRUE)
titanic <- titanic %>% mutate(age = replace(age, is.na(age), mean_age))
```


### Problem 4
**3:Lifeboat**
You’re interested in looking at the distribution of passengers in different lifeboats, but as we know, many passengers did not make it to a boat :-( This means that there are a lot of missing values in the boat column. Fill these empty slots with a dummy value e.g. the string 'None' or 'NA'

```{r}
 titanic <- titanic %>% mutate(boat = replace(boat, is.na(boat), "None"))
```


###Problem 5
**4:Cabin**
You notice that many passengers don’t have a cabin number associated with them.
Does it make sense to fill missing cabin numbers with a value?
What does a missing value here mean?
You have a hunch that the fact that the cabin number is missing might be a useful indicator of survival. Create a new column has_cabin_number which has 1 if there is a cabin number, and 0 otherwise.

```{r}
titanic <- titanic %>% mutate(cabin = replace(cabin, is.na(cabin), "NA"))
titanic <- titanic %>% mutate(has_cabin_number = ifelse(cabin == "NA", 0, 1))
```
Thank you.....
Finish by saving the ammended file to "titanic_refined.csv"

```{r}
write.csv(titanic, file = "titanic_refined.csv")